### 2021-05-21 Rmd/CP_eyetrackingR.Rmd
Added analysis section
Code for analysing model fit (growth curve) and bonferroni tests for specific comparisons of interest

---
title: "CognatePriming_EyetrackingR"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load packages
library("Matrix")
library("eyetrackingR")
library("tidyverse")
library("lme4")
library("ggplot2")
library("multcomp") # bonferroni

```

```{r prepare, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

# load data
processed_oxford <- readRDS(here("Results", "processed_oxford.rds")) %>%
  mutate(sample = "oxford")

processed <- bind_rows(processed_oxford)

```

## PREPARING DATA FOR ANALYSIS

```{r eyetrackingR, eval=TRUE, echo = FALSE}

processed_clean <- processed %>%
  filter(valid_trial == TRUE) %>%
  filter(valid_participant == TRUE) %>%
  mutate(trackloss_col = ifelse(valid_sample == TRUE, FALSE, TRUE)) # needed for eyetrackingR - TRUE means trackloss, FALSE means valid gaze

data <- make_eyetrackingr_data(processed_clean, 
                                      participant_column = "participant",
                                      trial_column = "trial_num",
                                      time_column = "time",
                                      trackloss_column = "trackloss_col",
                                      item_columns = c("prime", "target"),
                                      aoi_columns = c('aoi_target','aoi_distractor'), 
                                      treat_non_aoi_looks_as_missing = FALSE)

```

#### Aggregrate timebins + specify Target as AOI

```{r time_sequence, eval=TRUE, echo = FALSE}

# aggregate across trials within subjects in time analysis
response_time <- make_time_sequence_data(data, time_bin_size = 0.1, 
                                                predictor_columns = c("test_language", "sample", "trial_type", "age_group", "lp"),
                                                aois = "aoi_target")
#summary(response_time)

```

### Descriptives

```{r descriptive, eval=TRUE, echo = FALSE}

# Number of participants after cleaning
print("Final participants after data cleaning")

response_time %>%
  dplyr::select(participant, sample, age_group, lp) %>%
  unique() %>%
  group_by(sample, age_group, lp) %>%
  summarise(N = n())

```

## ANALYSIS
### Growth curve

```{r visualise only, eval=TRUE, echo = FALSE}

response_time_mono_21 <- response_time %>%
  filter(lp == "Monolingual" & age_group == "21 months")
response_time_mono_25 <- response_time %>%
  filter(lp == "Monolingual" & age_group == "25 months")
response_time_mono_30 <- response_time %>%
  filter(lp == "Monolingual" & age_group == "30 months")

# Visualise data from each group
# Monolingual

############ 21 months ##################

print("21 month monolingual data")

plot(response_time_mono_21, predictor_column = "trial_type", dv = "Prop") +
  theme_light() +
  xlab("Time from target & distractor image onset (s)") +
  ylab("Proportion of time looking at target") +
  ylim(c(0:1))

print("21 month monolingual data with model fit curve")

############ 25 months ##################

print("25 month monolingual data")

plot(response_time_mono_25, predictor_column = "trial_type", dv = "Prop") +
  theme_light() +
  xlab("Time from target & distractor image onset (s)") +
  ylab("Proportion of time looking at target") +
  ylim(c(0:1))

############ 30 months ##################

print("30 month monolingual data")

plot(response_time_mono_30, predictor_column = "trial_type", dv = "Prop") +
  theme_light() +
  xlab("Time from target & distractor image onset (s)") +
  ylab("Proportion of time looking at target") +
  ylim(c(0:1))

```

```{r analysis, eval=TRUE, echo = FALSE}

# Plot 21 month monolingual data with model fit curve
model_time_sequence_mono_21 <- lmer(Prop ~ trial_type*(ot1 + ot2 + ot3) + 
                                               (1 | participant), 
                                             data = response_time_mono_21, REML = FALSE)

plot(response_time_mono_21, predictor_column = "trial_type", dv = "Prop", model = model_time_sequence_mono_21) +
  theme_light() +
  xlab("Time from target & distractor image onset (s)") +
  ylab("Proportion of time looking at target") +
  ylim(c(0:1))

summary(model_time_sequence_mono_21)

# Post-hoc tests - only analyse significant predictors / interactions
#library(multcomp)
# To construct matrix, refer to summary()
contrast.matrix = rbind(
  "CP-NP" = c(0,-1,0, 0,0,0, 0,0,0,0,0,0), # ConditionCognate Prime(REFERENCE) - ConditionNon-Cognate Prime (Is there a significant difference between CP and NP?)
  "NP-UN" = c(0,1,-1, 0,0,0, 0,0, 0,0, 0,0), # ConditionNon-Cognate Prime - ConditionUnrelated Prime (Is there a significant difference between NP and UN?)
  "CP-NP_OT1" = c(0,0,0, 0,0,0, -1,0, 0,0, 0,0), # ConditionCognate Prime(REFERENCE) - ConditionNon-Cognate Prime (OT1, linear)
  "NP-UN_OT1" = c(0,0,0, 0,0,0, -1,1, 0,0, 0,0) # ConditionNon-Cognate Prime - ConditionUnrelated Prime (OT1, linear)
)
comps <- glht(model_time_sequence_mono_21, contrast.matrix )
summary(comps, test= adjusted("bonferroni"))

```