### 2022-01-04 presentations/2022-01-25_lacre/2022-01-25_lacre-abstract.Rmd
Update participants

---
title: "LACRE 2022: abstract"
author: "Gonzalo García-Castro"
date: "25/01/2022"
output: 
  word_document: 
    reference_docx: "manuscript_template.docx"
bibliography: "../../references.bib"
---

Previous studies have provided evidence that bilingual lexical access is language non-selective: recognising and producing words in one language activates lexical representations of words in the other language [e.g., @costa2000cognate; @thierry2007brain]. It has been suggested that this parallel activation is already present during toddlerhood [e.g., @von2012language; @jardak2019labels], but it is unclear how this it impacts the developing lexicon. We tested Catalan-Spanish bilinguals in a primed word recognition task, in which the auditory stimuli presented during the task were exclusively in participants' dominant language. We registered participants' looking preference for named (target) pictures as an indicator of word recognition. In primed trials, target pictures were preceded by the presentation of a silent picture whose label shared phonological onset with the target label. In non-primed trials, both labels had different onsets. We also manipulated the cognate status of the prime labels: some prime labels also shared phonological onset with their translation in participants' other language, while others did not. This resulted in three types of trials: trials in which prime and target labels shared phonological onset and the prime label was a cognate (cognate trials; e.g., *flower*/*flor*-*bird*, for an English-Spanish participant), trials where prime and target labels shared phonological onset but the prime label was a non-cognate (non-cognate trials; e.g., *chair*/*silla*-*bird*), and trials where prime and target labels did not share phonological onset (unrelated trials; *car*/*coche*-*bird*).

In line with previous studies, we expected participants to generate implicit labels for the prime pictures, which should interfere with target word recognition when both words share phonological onset [e.g.; @mani2010infant]. Under the hypothesis that bilingual participants would activate labels for the prime pictures in both languages, we further predicted  that interference over target word recognition would be stronger after cognate primes (labels from both languages interfere with target word recognition) than after non-cognate primes (only the label in the target language interferes with target word recognition). We tested bilinguals (and same-aged monolingual controls) at three age points (21, 25, and 30 months) to investigate how any cross-language priming effect emerged or changed across these ages, capitalising on the potential role of vocabulary size. Preliminary data suggest that bilinguals were almost insensitive to the phonological similarity between the prime and the spoken target word, or to the cognate status of the prime label. Surprisingly, monolinguals' preference for the target picture at 25 was stronger after non-cognate primes, compared to cognate primes or unrelated primes, and also after cognate and non-cognate primes compared to unrelated primes. We present this preliminary data along with exploratory analyses addressing the puzzling pattern of results we observed.

![](img/gca.png)