### 2023-11-03 manuscript/_appendix/_vocabulary-validity.qmd

## Appendix D: vocabulary checklist validity

```{r fig-validity-target}
gaze_tmp <- gaze |> 
	filter(phase=="Target-Distractor",
		   between(timestamp, time_subset[1], time_subset[2])) |> 
	select(session_id, trial, phase, timestamp,
		   is_gaze_target, is_gaze_distractor, trial_type) |> 
	mutate(condition = recode_condition(trial_type))

vocabulary_tmp <- rename_with(vocabulary,  
							  \(x) gsub("_prop", "", paste0("voc_", x)), 
							  matches("_prop"))

participants_tmp <- participants |> 
	select(child_id, session_id, age_group, age, lp)

attrition_trials_tmp <- get_attrition_trials(
	participants = participants,
	vocabulary = vocabulary,
	vocabulary_by = c("target"),
	aoi_coords = aoi_coords,
	gaze = gaze,
	min_looking = c(prime = 0.75,
					test = 1.00,
					test_each = 0.00,
					test_any = 0.1)) |> 
	select(session_id, trial, samples, is_valid_vocab_all) |> 
	unnest_wider(is_valid_vocab_all) |> 
	rename_with(\(x) gsub("is_valid_vocab_", "", x))

attrition_participants_tmp <- attrition_participants |> 
	filter(is_valid_participant) |> 
	select(session_id)

data <- gaze_tmp |>
	inner_join(attrition_trials_tmp, by = join_by(session_id, trial)) |> 
	inner_join(attrition_participants_tmp, by = join_by(session_id)) |> 
	mutate(condition = recode_condition(trial_type)) |> 
	select(session_id, trial, condition, is_gaze_target, is_gaze_distractor,
		   is_valid_vocab_all) |> 
	# aggregated across trials by participant, time bin and condition
	# see Chow et al. (2018)
	summarise(sum_target = sum(is_gaze_target, na.rm = TRUE),
			  sum_distractor = sum(is_gaze_distractor, na.rm = TRUE),
			  .nsamples = sum(is_gaze_target | is_gaze_distractor, na.rm = TRUE),
			  .ntrials = length(unique(trial)),
			  .by = c(session_id, condition, is_valid_vocab_all)) |> 
	# empirical logit with adjustment
	# see Barr et al. (2008)
	mutate(.nsamples = sum_target + sum_distractor,
		   .prop = ifelse(.nsamples==0, 0, sum_target / .nsamples),
		   .elog = log((sum_target + .5)/(sum_distractor + .5))) |> 
	rename(.sum = sum_target) |> 
	arrange(desc(session_id), condition) |> 
	inner_join(participants_tmp, by = join_by(session_id)) |> 
	inner_join(vocabulary_tmp, by = join_by(child_id, session_id)) |> 
	mutate(across(c(.nsamples), as.integer),
		   across(c(child_id, session_id, lp, condition, age_group), as.factor),
		   across(c(age, voc_l1),
		   	   \(x) scale(x, scale = TRUE)[, 1],
		   	   .names = "{.col}_std")) |> 
	select(child_id, session_id, age_group, age, voc_l1, voc_total, lp,
		   condition,  .sum, .prop, .elog, .nsamples,
		   matches("std"), is_valid_vocab_all)

# set a priori contrasts
contrasts(data$condition) <- cbind(c(-0.5, 0.5, 0),
								   c(0, -0.5, 0.5))
contrasts(data$lp) <- cbind(c(-0.5, 0.25, 0.25),
							c(0, -0.5, 0.5))
contrasts(data$age_group) <- cbind(c(-0.5, 0.5, 0),
								   c(0, -0.5, 0.5))

data |> 
	unnest_wider(is_valid_vocab_all) |>
	rename_with(\(x) gsub("is_valid_vocab_", "", x)) |> 
	select(session_id, lp, .elog, target) |> 
	arrange(session_id, target) |> 
	summarise(.elog_mean = mean(.elog),
			  .elog_sd = sd(.elog),
			  n = n(),
			  .by = c(lp, target)) |> 
	mutate(target = ifelse(target, "Yes", "No"),
		   .elog_se = .elog_sd/sqrt(n),
		   .elog_lower = .elog_mean - .elog_se*1.96,
		   .elog_upper = .elog_mean + .elog_se*1.96,
	) |> 
	ggplot(aes(lp, .elog_mean, 
			   ymin = .elog_lower,
			   ymax = .elog_upper,
			   colour = target)) +
	facet_wrap(~lp, ncol = 1, scales = "free_x") +
	geom_errorbar(width = 0.2, linewidth = 3/4, 
				  position = position_dodge(width = 0.25)) +
	geom_point(size = 2,
			   position = position_dodge(width = 0.25)) +
	geom_hline(yintercept = 0,
			   linetype = "dashed") +
	labs(x = "Condition",
		 y = "P(Target looking)", 
		 colour = "Knows the target word") +
	theme(legend.position = "none",
		  panel.grid.major.y = element_line(colour = "grey",
		  								  linetype = "dotted"))

get_data(gaze = gaze,
		 participants = participants,
		 stimuli = stimuli,
		 vocabulary = vocabulary,
		 attrition_trials = attrition_trials,
		 attrition_participants = attrition_participants,
		 time_subset = c(0, 2),
		 contrast = "related",
		 is_valid_vocab_all,
		 clean = FALSE) |> 
	unnest_wider(is_valid_vocab_all) |>
	select(session_id, age, lp, timebin, .elog, target) |> 
	arrange(session_id, target) |> 
	summarise(.elog_mean = mean(.elog),
			  .elog_sd = sd(.elog),
			  n = n(),
			  .by = c(timebin, lp, target)) |> 
	mutate(target = ifelse(target, "Yes", "No"),
		   .elog_se = .elog_sd/sqrt(n),
		   .elog_lower = .elog_mean - .elog_se*1.96,
		   .elog_upper = .elog_mean + .elog_se*1.96,
	) |> 
	ggplot(aes(timebin, .elog_mean,
			   ymin = .elog_lower,
			   ymax = .elog_upper,
			   colour = target,
			   fill = target)) +
	facet_wrap(~lp, ncol = 1, scales = "free_x") +
	geom_line(linewidth = 1/2) +
	geom_ribbon(alpha = 1/2,
				colour = NA) +
	geom_point(size = 2) +
	geom_hline(yintercept = 0,
			   linetype = "dashed") +
	labs(x = "Condition",
		 y = "P(Target looking)", 
		 fill = "Knows the target word",
		 colour = "Knows the target word") +
	theme(legend.position = "top",
		  panel.grid.major.y = element_line(colour = "grey",
		  								  linetype = "dotted"),
		  axis.title.y = element_blank(),
		  axis.line.x = element_line()) +

	plot_layout(ncol = 2,
				widths = c(0.2, 0.8)) &
	scale_y_continuous(limits = c(-0.5, 1.25))
```